### 2020-10-20 Ant Matrix.Rmd
initial commit of Ali's `Shared Code' folder contents.

---
title: "Ant Transition Matrix"
author: "Ali Campbell"
date: "9/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
```

```{r}
#import the data
cactus <- read.csv("/Users/alicampbell/Desktop/cholla.csv", header = TRUE)
#str(cactus)
```

##Data Filtering
```{r}
## function for the volume of a cone

volume <- function(h, w, p){
  (1/3)*pi*h*(((w + p)/2)/2)^2
}
invlogit <- function(x){exp(x)/(1+exp(x))}

#Create volume columns
cactus$volume_t <- volume(cactus$Height_t,cactus$Width_t, cactus$Perp_t)
cactus$volume_t1 <- volume(cactus$Height_t1,cactus$Width_t1, cactus$Perp_t1)

#hist(log(cactus$volume_t))
```

```{r}
##Year T
cactus <- data.frame(lapply(cactus,      # Convert data with toupper function
                              function(variables) {
                                if (is.character(variables)) {
                                  return(toupper(variables))
                                } else {
                                  return(variables)
                                }
                              }),
                       stringsAsFactors = FALSE)
cactus_new <- as.data.frame(gsub("LIOM", "L",cactus$Antsp_t))
cactus_new <- as.data.frame(gsub("CREM", "C",cactus$Antsp_t))
cactus_new <- as.data.frame(gsub("FOR", "F",cactus$Antsp_t))

##Year T1
cactus <- data.frame(lapply(cactus,      # Convert data with toupper function
                              function(variables) {
                                if (is.character(variables)) {
                                  return(toupper(variables))
                                } else {
                                  return(variables)
                                }
                              }),
                       stringsAsFactors = FALSE)
cactus_new <- as.data.frame(gsub("LIOM", "L",cactus$Antsp_t1))
cactus_new <- as.data.frame(gsub("CREM", "C",cactus$Antsp_t1))
cactus_new <- as.data.frame(gsub("FOR", "F",cactus$Antsp_t1))

#unique(cactus$Antsp_t)
```

```{r}
#Bin all ant species in year t to a value 1 - 5 without taking away the species denotation
cactus$bin1 <- 0 #blank column 
for(i in seq_len(nrow(cactus))){
  if(cactus$Antsp_t[i] == "LIOM"){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "L"){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "LIOM "){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "CREM"){cactus$bin1[i] <- 2}
  else if(cactus$Antsp_t[i] == "C"){cactus$bin1[i] <- 2}
  else if(cactus$Antsp_t[i] == "FOR"){cactus$bin1[i] <- 3}
  else if(cactus$Antsp_t[i] == "F"){cactus$bin1[i] <- 3}
  else if(cactus$Antsp_t[i] == "VACANT"){cactus$bin1[i] <- 4}
  else if(cactus$Antsp_t[i] == ""){cactus$bin1[i] <- 4}
  else{cactus$bin1[i] <- 5}
}
#Bin all ant species in yeay t1 to a value 1 - 5 without taking away species denotation
cactus$bin2 <- 0 #blank column 
for(i in seq_len(nrow(cactus))){
  if(cactus$Antsp_t1[i] == "LIOM"){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "L"){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "LIOM "){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "CREM"){cactus$bin2[i] <- 2}
  else if(cactus$Antsp_t1[i] == "C"){cactus$bin2[i] <- 2}
  else if(cactus$Antsp_t1[i] == "FOR"){cactus$bin2[i] <- 3}
  else if(cactus$Antsp_t1[i] == "F"){cactus$bin2[i] <- 3}
  else if(cactus$Antsp_t1[i] == "VACANT"){cactus$bin2[i] <- 4}
  else if(cactus$Antsp_t1[i] == ""){cactus$bin2[i] <- 4}
  else{cactus$bin2[i] <- 5}
}
```

##Ant Transition Matrix
#### Hard-Coded Matrix
```{r}
transMat <- as.data.frame(prop.table(with(cactus, table(cactus$bin1, cactus$bin2)), 2))

Transitions <- matrix(c(0.66064710	, 0.04138450	, 0.01504891	, 0.27012792	, 0.01279157 ,	0.14760148	, 0.29889299	, 0.01660517	, 0.50553506	, 0.03136531	, 0.27131783	, 0.09302326	, 0.06201550	, 0.50387597	, 0.06976744	, 0.21302405	, 0.06086401	, 0.01532241	, 0.69035965	, 0.02042988	, 0.27950311	, 0.21118012	, 0.03105590	, 0.43478261	, 0.04347826),ncol = 5, nrow = 5, byrow = FALSE)
#Transitions
#colSums(Transitions)
```

#### Not Hard-Coded matrix
```{r}
# If you have for example a matrix like this
m <- matrix(c(1:25), ncol=5, nrow=5)

for(i in 1:5){
  for(j in 1:5){
    m[i,j] <- transMat[i,3]
  }
}
```

##Plot Survival by Volume probability Data
```{r}
plot(log(cactus$volume_t),cactus$Survival_t1)

survival_plot <- cactus %>% 
  mutate(size_bin = cut_interval(log(volume_t),10)) %>%
  group_by(size_bin) %>%
  summarise(mean_size = mean(log(volume_t),na.rm=T),
    surv = mean(Survival_t1,na.rm=T))

years <- unique(cactus$Year_t)
n_years <- length(years)
points(survival_plot$mean_size,survival_plot$surv,pch=16,cex=2,col="red")

```

## Vital Rate Plots
#### Survival Rate

Three models were considered here. First A simple survival vs log(volume) model. This had an AIC value of 2380.825.  The second included a second explanatory variable of the ant species (as an independent variable). This had an AIC vaue of 2474.975/ The third included the ant species variable as an interaction term with the volume. This had an AIC value of 2378.601, which is a difference greater than 2, and therefore is a satistically supported improvement. 
```{r}
surv_model <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus)
surv_model_ant <- glm(Survival_t1 ~ 0 + log(volume_t) + as.factor(bin1),
                  family="binomial",
                  data=cactus)
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)
```

```{r}
#plot the basic surv_model with plot
plot(log(cactus$volume_t),cactus$Survival_t1)
size_dummy <- seq(min(log(cactus$volume_t),na.rm=T),max(log(cactus$volume_t),na.rm=T),0.01)
lines(size_dummy,invlogit(coef(surv_model)[1] + coef(surv_model)[2]*size_dummy), lwd=3, col="red")
#now with ggplot
ggplot(cactus, aes(y=cactus$Survival_t1, x=log(cactus$volume_t))) +  geom_point() +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

This includes an ant interaction term
```{r}
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)

#anova(surv_model_ant_interact, test = "Chisq")
#summary(surv_model_ant_interact)
```

```{r}
ggplot(cactus, aes(y=Survival_t1, x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.1) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

#### Flowering Model
```{r}
flo_model <- glm(Goodbuds_t > 0 ~ log(volume_t),
                  family="binomial",
                  data=cactus)

cactus$Flow <- cactus$Goodbuds_t > 0 #blank column 
for(i in 1:8189){
  if(cactus$TotFlowerbuds_t[i] == NA){
    cactus$Flow[i] <- sum(cactus$ABFlowerbuds_t[i]+cactus$Goodbuds_t[i])
  }
  else if(cactus$TotFlowerbuds_t[i] == 0){
    cactus$Flow[i] <- 0
  }
  else if(cactus$TotFlowerbuds_t[i] >0){
    cactus$Flow[i] <- 1
  }
}

cactus$Flow
ggplot(cactus, aes(y=Flow, x=log(volume_t))) +  geom_point() +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE)

plot(log(cactus$volume_t),cactus$Flow)
```

#### Growth Model
```{r}
gro_model_ant_interact <- glm(log(volume_t1) ~ 0 + log(volume_t) * as.factor(bin1),
                  data=cactus)

ggplot(cactus, aes(y=log(volume_t1), x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', se = FALSE) 
```
## Ant Transition Model
```{r}
ant_trans_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1),family = binomial,data=cactus)
ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

```{r}
ant_trans_size_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) + log(volume_t),family = binomial,data=cactus)
ant_trans_size_model2 <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) * log(volume_t),family = binomial,data=cactus)
#AIC(ant_trans_model)
#AIC(ant_trans_size_model)
#AIC(ant_trans_size_model2)
##Therefore we can see that the interaction transition model is the best. And that Size is an important factor.
summary(ant_trans_size_model2)
ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1)*log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
#model.matrix(ant_trans_size_model2)
```

## Multinomial Models

```{r}
#Will be ant vs size
x1 <- factor(as.factor(cactus$bin1), levels = c("5","1","2","3","4"))
m1 <- multinom(as.factor(cactus$bin1) ~ 0+ 1, data = cactus)
m2 <- multinom(x1 ~ 1, data = cactus)
z1<- summary(m1)$coefficients/summary(m1)$standard.errors
z2<- summary(m2)$coefficients/summary(m2)$standard.errors
summary(m1)
summary(m2)
#coef(m1)
#coef(m2)
p1 <- (1 - pnorm(abs(z1), 0, 1)) * 2
p2 <- (1 - pnorm(abs(z2), 0, 1)) * 2
ggplot(cactus, aes(x = cactus$bin2, y = cactus$bin1, colour = cactus$bin1)) + geom_line() 
#model.matrix(m1)s
```

```{r}
#Will be ant vs ant
m2 <- multinom(as.factor(bin1) ~ 0+as.factor(bin2), data = cactus)
z<- summary(m2)$coefficients/summary(m2)$standard.errors
#summary(m2)
p <- (1 - pnorm(abs(z), 0, 1)) * 2
ggplot(cactus, aes(x = bin2, y = bin1, colour = bin1)) + geom_line()
#model.matrix(m2)
```