### 2020-10-20 Survival Graphs Binned.Rmd

---
title: "Survival Graphs Binned"
author: "Ali Campbell"
date: "9/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load packages
library(tidyverse)
library(lme4)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
```

```{r}
#import the data
cactus <- read.csv("/Users/alicampbell/Desktop/cholla.csv", header = TRUE)
#str(cactus)
```

##Filter Ant Species to standardized
```{r}
cactus <- data.frame(lapply(cactus,      # Convert data with toupper function
                              function(variables) {
                                if (is.character(variables)) {
                                  return(toupper(variables))
                                } else {
                                  return(variables)
                                }
                              }),
                       stringsAsFactors = FALSE)

##for loop to change them
#Bin all ant species in year t to a value 1 - 5 without taking away the species denotation
cactus$bin1 <- 0 #blank column 
for(i in seq_len(nrow(cactus))){
  if(cactus$Antsp_t[i] == "LIOM"){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "L"){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "LIOM "){cactus$bin1[i] <- 1}
  else if(cactus$Antsp_t[i] == "CREM"){cactus$bin1[i] <- 2}
  else if(cactus$Antsp_t[i] == "C"){cactus$bin1[i] <- 2}
  else if(cactus$Antsp_t[i] == "FOR"){cactus$bin1[i] <- 3}
  else if(cactus$Antsp_t[i] == "F"){cactus$bin1[i] <- 3}
  else if(cactus$Antsp_t[i] == "VACANT"){cactus$bin1[i] <- 4}
  else if(cactus$Antsp_t[i] == ""){cactus$bin1[i] <- 4}
  else{cactus$bin1[i] <- 5}
}
#Bin all ant species in yeay t1 to a value 1 - 5 without taking away species denotation
cactus$bin2 <- 0 #blank column 
for(i in seq_len(nrow(cactus))){
  if(cactus$Antsp_t1[i] == "LIOM"){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "L"){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "LIOM "){cactus$bin2[i] <- 1}
  else if(cactus$Antsp_t1[i] == "CREM"){cactus$bin2[i] <- 2}
  else if(cactus$Antsp_t1[i] == "C"){cactus$bin2[i] <- 2}
  else if(cactus$Antsp_t1[i] == "FOR"){cactus$bin2[i] <- 3}
  else if(cactus$Antsp_t1[i] == "F"){cactus$bin2[i] <- 3}
  else if(cactus$Antsp_t1[i] == "VACANT"){cactus$bin2[i] <- 4}
  else if(cactus$Antsp_t1[i] == ""){cactus$bin2[i] <- 4}
  else{cactus$bin2[i] <- 5}
}
```

##Create Volume
```{r}
## function for the volume of a cone

volume <- function(h, w, p){
  (1/3)*pi*h*(((w + p)/2)/2)^2
}
invlogit <- function(x){exp(x)/(1+exp(x))}

#Create volume columns
cactus$volume_t <- volume(cactus$Height_t,cactus$Width_t, cactus$Perp_t)
cactus$volume_t1 <- volume(cactus$Height_t1,cactus$Width_t1, cactus$Perp_t1)

hist(log(cactus$volume_t))
```

##Plot Survival by Volume probability
```{r}
plot(log(cactus$volume_t),cactus$Survival_t1)

survival_plot <- cactus %>% 
  mutate(size_bin = cut_interval(log(volume_t),10)) %>%
  group_by(size_bin) %>%
  summarise(mean_size = mean(log(volume_t),na.rm=T),
    surv = mean(Survival_t1,na.rm=T))

years <- unique(cactus$Year_t)
n_years <- length(years)
points(survival_plot$mean_size,survival_plot$surv,pch=16,cex=2,col="red")

```

##Fit statistical Binom Model Tier 1
```{r}
#This is the simplest possible model
##The simplest possible model for size dep model
surv_model <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus)
surv_model_ant <- glm(Survival_t1 ~ 0 + log(volume_t) + as.factor(bin1),
                  family="binomial",
                  data=cactus)
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)
plot(log(cactus$volume_t),cactus$Survival_t1)
size_dummy <- seq(min(log(cactus$volume_t),na.rm=T),max(log(cactus$volume_t),na.rm=T),0.01)
lines(size_dummy,invlogit(coef(surv_model)[1] + coef(surv_model)[2]*size_dummy), lwd=3, col="red")
AIC(surv_model_ant)
AIC(surv_model)
AIC(surv_model_ant_interact)
#Lower is better for AIC, providing statistical support for the ant model
#AIC suggests that there is a size ant interaction
#AIC difference of 2 is a statistically supported improvement
```

```{r}
summary(surv_model_ant_interact)
b0 <- coef(surv_model_ant_interact)[1]
b1 <- coef(surv_model_ant_interact)[2]
b2 <- coef(surv_model_ant_interact)[3]
b3 <- coef(surv_model_ant_interact)[4]
b4 <- coef(surv_model_ant_interact)[5]
b5 <- coef(surv_model_ant_interact)[6]
b6 <- coef(surv_model_ant_interact)[7]
b7 <- coef(surv_model_ant_interact)[8]
b8 <- coef(surv_model_ant_interact)[9]
b9 <- coef(surv_model_ant_interact)[10]

```

```{r}
vol_range <- seq(from=min(log(na.omit(cactus$volume_t))), to=max(log(na.omit(cactus$volume_t))), by=.01)
mean
1logits <- 0 + b0*vol_range + 
```

```{r}
#plot(surv_model)
plot(log(cactus$volume_t),cactus$Survival_t1)
curve(exp(b0+b1*x)/(1+exp(b0+b1*x)),add=T)

ggplot(cactus, aes(y=cactus$Survival_t1, x=log(cactus$volume_t))) +  geom_point() +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

```{r}
flo_model <- glm(TotFlowerbuds_t > 0 ~ log(volume_t),
                  family="binomial",
                  data=cactus)
flo_model_ant <- glm(TotFlowerbuds_t > 0 ~ 0 + log(volume_t) + as.factor(bin1),
                  family="binomial",
                  data=cactus)
flo_model_ant_interact <- glm(TotFlowerbuds_t > 0 ~ 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)

AIC(flo_model)
AIC(flo_model_ant)
AIC(flo_model_ant_interact)

##In this case the ant models are not better by AIC
```

```{r}
summary(flo_model)
b0=coef(flo_model)[1]
b1=coef(flo_model)[2]
plot(log(cactus$volume_t1),cactus$TotFlowerbuds_t)
curve(exp(b0+b1*x)/(1+exp(b0+b1*x)),add=T)
```

```{r}
gro_model <- glm(log(volume_t1) ~ log(volume_t),
                  data=cactus)
gro_model_ant <- glm(log(volume_t1) ~ 0 + log(volume_t) + as.factor(bin1),
                  data=cactus)
gro_model_ant_interact <- glm(log(volume_t1) ~ 0 + log(volume_t) * as.factor(bin1),
                  data=cactus)

AIC(gro_model)
AIC(gro_model_ant)
AIC(gro_model_ant_interact)
##The ant interaction model proves the best again
```

```{r}
ggplot(cactus, aes(y=Survival_t1, x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.1) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

#Plot survival probability by volume, grouped by ant species (the bins)
```{r}
survival_plot <- cactus %>% 
  mutate(size_bin = cut_interval(log(volume_t),10)) %>% 
  group_by(size_bin, cactus$bin1) %>% 
  summarise(mean_size = mean(log(volume_t),na.rm=T),
    surv = mean(Survival_t1,na.rm=T))

plot(log(cactus$volume_t),cactus$Survival_t1)
points(survival_plot$mean_size,survival_plot$surv,pch=16,cex=1, col=1:length(cactus$bin1))
legend(12,1,unique(cactus$bin1),col=1:length(cactus$bin1),pch=1)
```

##Fit Statistical Binom Model Tier 2
####Independent
```{r}
surv_model_2 <- glmer(Survival_t1 ~ log(volume_t) + (1|Year_t) + (1|Plot),
                  family="binomial",
                  data=cactus)
plot(log(cactus$volume_t),cactus$Survival_t1)
size_dummy_2 <- seq(min(log(cactus$volume_t),na.rm=T),max(log(cactus$volume_t),na.rm=T),0.01)
lines(size_dummy_2,invlogit(fixef(surv_model_2)[1] + fixef(surv_model_2)[2]*size_dummy_2), lwd=3, col="red")
lines(size_dummy,invlogit(coef(surv_model)[1] + coef(surv_model)[2]*size_dummy), lwd=3, col="blue")
```

####Interaction
```{r}
surv_model_2 <- glmer(Survival_t1 ~ log(volume_t) + ((log(volume_t)|Year_t) + (log(volume_t)|Plot)),
                  family="binomial",
                  data=cactus)
plot(log(cactus$volume_t),cactus$Survival_t1)
size_dummy_2 <- seq(min(log(cactus$volume_t),na.rm=T),max(log(cactus$volume_t),na.rm=T),0.01)
lines(size_dummy_2,invlogit(fixef(surv_model_2)[1] + fixef(surv_model_2)[2]*size_dummy_2), lwd=3, col="red")
lines(size_dummy,invlogit(coef(surv_model)[1] + coef(surv_model)[2]*size_dummy), lwd=3, col="blue")
```

```{r}
cactus_split <- split(cactus,cactus$bin1)
surv_model_bin1 <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus_split$`1`)
size_dummy <- seq(min(log(cactus_split$`1`$volume_t),na.rm=T),max(log(cactus_split$`1`$volume_t),na.rm=T),0.01)
surv_model_bin2 <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus_split$`2`)
plot(log(cactus_split$`2`$volume_t),cactus_split$`2`$Survival_t1, xlab = "log(Volume)", ylab = "Survival")
surv_model_bin3 <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus_split$`3`)
surv_model_bin4 <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus_split$`4`)
surv_model_bin5 <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus_split$`5`)

size_dummy <- seq(min(log(cactus_split$`2`$volume_t),na.rm=T),max(log(cactus_split$`2`$volume_t),na.rm=T),0.01)
lines(size_dummy,invlogit(coef(surv_model_bin1)[1] + coef(surv_model_bin1)[2]*size_dummy), lwd=3, col="red")
lines(size_dummy,invlogit(coef(surv_model_bin2)[1] + coef(surv_model_bin2)[2]*size_dummy), lwd=3, col="blue")
lines(size_dummy,invlogit(coef(surv_model_bin3)[1] + coef(surv_model_bin3)[2]*size_dummy), lwd=3, col="green")
lines(size_dummy,invlogit(coef(surv_model_bin4)[1] + coef(surv_model_bin4)[2]*size_dummy), lwd=3, col="yellow")
lines(size_dummy,invlogit(coef(surv_model_bin5)[1] + coef(surv_model_bin5)[2]*size_dummy), lwd=3, col="orange")
legend(11.5,.4,legend = c("Liom","Crem","For","Vacant","Other"),col = c("red","blue","green","yellow","orange"),lty = 1,cex = .8)
```

```{r}
AIC(surv_model_bin1)
AIC(surv_model_bin2)
```