### 2020-10-21 Ant Matrix_TOM.Rmd
progress on new file

---
title: "Ant Transition Matrix"
author: "Ali Campbell"
date: "9/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
```

```{r read_data}
#import the data
cactus <- read.csv("cholla_demography_20042019.csv", header = TRUE)
#str(cactus) ##<- problem: antcount is a factor
#levels(cactus$Antcount_t);levels(cactus$Antcount_t1)
## drop the offending rows -- this is a 2019 data entry problem
cactus <- cactus[-c(which(cactus$Antcount_t=="2-Jan"),which(cactus$Antcount_t1=="2-Jan")),]
cactus$antcount_t <- as.numeric(as.character(cactus$Antcount_t))
cactus$antcount_t1 <- as.numeric(as.character(cactus$Antcount_t1))
```

##Data Filtering
```{r volume}
## function for the volume of a cone

volume <- function(h, w, p){
  (1/3)*pi*h*(((w + p)/2)/2)^2
}
invlogit <- function(x){exp(x)/(1+exp(x))}

#Create volume columns
cactus$volume_t <- volume(cactus$Height_t,cactus$Width_t, cactus$Perp_t)
cactus$volume_t1 <- volume(cactus$Height_t1,cactus$Width_t1, cactus$Perp_t1)

#hist(log(cactus$volume_t))
```

```{r year_t_ants}
## assign ant counts of zero as vacant
cactus$Antsp_t[cactus$antcount_t==0] <- "vacant"
# here are the ordered levels of the current variable
Antsp_t_levels <- levels(cactus$Antsp_t)
# here is how I would like to collapse these into fewer bins -- most will be "other"
ant_t_levels <- rep("other",times=length(Antsp_t_levels))
## crem levels - elements 5,8,9
ant_t_levels[c(5,8,9)] <- "crem"
## liom levels - elements 15,18,19,20,21
ant_t_levels[c(15,18,19,20,21)] <- "liom"
## forelius levels - 12,13
ant_t_levels[c(12,13)] <- "for"
## vacant - 31
ant_t_levels[31] <- "vacant"
## create new variable merging levels as above
cactus$ant_t <- factor(cactus$Antsp_t,levels=Antsp_t_levels,labels=ant_t_levels)
summary(cactus$ant_t) ## there are too many "others". check these out.
filter(cactus,ant_t=="other")
## first, there are no ant data from 2007:
cactus$Antcount_t[cactus$Year_t==2007]
#so give these observations NA for ant status:
cactus$ant_t[cactus$Year_t==2007] <- NA
# also, if a plant was new in year t+1 it's year t ant status should be NA
cactus$ant_t[cactus$Newplant==1] <- NA
# plots 7 and 8 were added in 2011, so their year_t==2010 ant status should be NA
cactus$ant_t[cactus$Year_t==2010 & cactus$Plot==7] <- NA
cactus$ant_t[cactus$Year_t==2010 & cactus$Plot==8] <- NA
# check to see how many others this leaves -- still a lot
summary(cactus$ant_t)
## another 2019 data entry problem: new plants were not given Newplant==1 and their Survival is 0 (should be NA)
## for now here is my workaround: year t ==2018, height_t ==NA, height_t1 !=NA
cactus$ant_t[cactus$Year_t==2018 & is.na(cactus$Height_t) & !is.na(cactus$Height_t1)] <- NA
## finally there are these plants with Antsp_t=="" for all kinds of reasons but bottom line is that we don't have ant status
cactus$ant_t[cactus$Antsp_t==""]<-NA
## inspect the "others" visually
cactus %>% filter(ant_t=="other") %>% select(Antsp_t,ant_t)
## I am satisfied that everything assigned other is correctly assigned

```

```{r ant_time_series}
## plot a time series of ant frequency by year
ant_freq_ts <- cactus %>% 
  select(Year_t,ant_t) %>% 
  drop_na() %>% 
  group_by(Year_t,ant_t) %>% 
  summarise(n = n()) %>% 
  mutate(freq = n / sum(n))

plot(ant_freq_ts$Year_t[ant_freq_ts$ant_t=="vacant"],ant_freq_ts$freq[ant_freq_ts$ant_t=="vacant"],type="b",ylim=c(0,1),pch=16,cex=1.5,
     xlab="Year",ylab="Frequency")
legend("topleft",legend=c("Vacant","Liom.","Crem.","For.","Other"),pch=16,col=c("black","blue","red","green","pink"))
points(ant_freq_ts$Year_t[ant_freq_ts$ant_t=="liom"],ant_freq_ts$freq[ant_freq_ts$ant_t=="liom"],type="b",pch=16,cex=1.5,col="blue")
points(ant_freq_ts$Year_t[ant_freq_ts$ant_t=="crem"],ant_freq_ts$freq[ant_freq_ts$ant_t=="crem"],type="b",pch=16,cex=1.5,col="red")
points(ant_freq_ts$Year_t[ant_freq_ts$ant_t=="for"],ant_freq_ts$freq[ant_freq_ts$ant_t=="for"],type="b",pch=16,cex=1.5,col="green")
points(ant_freq_ts$Year_t[ant_freq_ts$ant_t=="other"],ant_freq_ts$freq[ant_freq_ts$ant_t=="other"],type="b",pch=16,cex=1.5,col="pink")

```

```{r year_t1_ants}
## assign ant counts of zero as vacant
cactus$Antsp_t1[cactus$antcount_t1==0] <- "vacant"
## repeat for t1 -- these indices are different
Antsp_t1_levels <- levels(cactus$Antsp_t1)
# here is how I would like to collapse these into fewer bins -- most will be "other"
ant_t1_levels <- rep("other",times=length(Antsp_t1_levels))
## crem levels 
ant_t1_levels[c(8,9,10)] <- "crem"
## liom levels 
ant_t1_levels[c(19,20,21,22)] <- "liom"
## forelius levels 
ant_t1_levels[c(13,14,15)] <- "for"
## vacant 
ant_t1_levels[32] <- "vacant"
## create new variable merging levels as above
cactus$ant_t1 <- factor(cactus$Antsp_t1,levels=Antsp_t1_levels,labels=ant_t1_levels)
summary(cactus$ant_t1) ## there are too many "others" -- same deal as above
## first, there are no ant data from 2007 so give these observations NA for ant status:
cactus$ant_t1[cactus$Year_t1==2007] <- NA
# check to see how many others this leaves -- still a lot
summary(cactus$ant_t1)
## another 2019 data entry problem: new plants were not given Newplant==1 and their Survival is 0 (should be NA)
## for now here is my workaround: year t ==2018, height_t ==NA, height_t1 !=NA
cactus$ant_t[cactus$Year_t==2018 & is.na(cactus$Height_t) & !is.na(cactus$Height_t1)] <- NA
## finally there are these plants with Antsp_t=="" for all kinds of reasons but bottom line is that we don't have ant status
cactus$ant_t[cactus$Antsp_t==""]<-NA
## inspect the "others" visually
cactus %>% filter(ant_t=="other") %>% select(Antsp_t,ant_t)
## I am satisfied that everything assigned other is correctly assigned

```

##Ant Transition Matrix
#### Hard-Coded Matrix
```{r}
transMat <- as.data.frame(prop.table(with(cactus, table(cactus$bin1, cactus$bin2)), 2))

Transitions <- matrix(c(0.66064710	, 0.04138450	, 0.01504891	, 0.27012792	, 0.01279157 ,	0.14760148	, 0.29889299	, 0.01660517	, 0.50553506	, 0.03136531	, 0.27131783	, 0.09302326	, 0.06201550	, 0.50387597	, 0.06976744	, 0.21302405	, 0.06086401	, 0.01532241	, 0.69035965	, 0.02042988	, 0.27950311	, 0.21118012	, 0.03105590	, 0.43478261	, 0.04347826),ncol = 5, nrow = 5, byrow = FALSE)
#Transitions
#colSums(Transitions)
```

#### Not Hard-Coded matrix
```{r}
# If you have for example a matrix like this
m <- matrix(c(1:25), ncol=5, nrow=5)

for(i in 1:5){
  for(j in 1:5){
    m[i,j] <- transMat[i,3]
  }
}
```

##Plot Survival by Volume probability Data
```{r}
plot(log(cactus$volume_t),cactus$Survival_t1)

survival_plot <- cactus %>% 
  mutate(size_bin = cut_interval(log(volume_t),10)) %>%
  group_by(size_bin) %>%
  summarise(mean_size = mean(log(volume_t),na.rm=T),
    surv = mean(Survival_t1,na.rm=T))

years <- unique(cactus$Year_t)
n_years <- length(years)
points(survival_plot$mean_size,survival_plot$surv,pch=16,cex=2,col="red")

```

## Vital Rate Plots
#### Survival Rate

Three models were considered here. First A simple survival vs log(volume) model. This had an AIC value of 2380.825.  The second included a second explanatory variable of the ant species (as an independent variable). This had an AIC vaue of 2474.975/ The third included the ant species variable as an interaction term with the volume. This had an AIC value of 2378.601, which is a difference greater than 2, and therefore is a satistically supported improvement. 
```{r}
surv_model <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus)
surv_model_ant <- glm(Survival_t1 ~ 0 + log(volume_t) + as.factor(bin1),
                  family="binomial",
                  data=cactus)
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)
```

```{r}
#plot the basic surv_model with plot
plot(log(cactus$volume_t),cactus$Survival_t1)
size_dummy <- seq(min(log(cactus$volume_t),na.rm=T),max(log(cactus$volume_t),na.rm=T),0.01)
lines(size_dummy,invlogit(coef(surv_model)[1] + coef(surv_model)[2]*size_dummy), lwd=3, col="red")
#now with ggplot
ggplot(cactus, aes(y=cactus$Survival_t1, x=log(cactus$volume_t))) +  geom_point() +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

This includes an ant interaction term
```{r}
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)

#anova(surv_model_ant_interact, test = "Chisq")
#summary(surv_model_ant_interact)
```

```{r}
ggplot(cactus, aes(y=Survival_t1, x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.1) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

#### Flowering Model
```{r}
flo_model <- glm(Goodbuds_t > 0 ~ log(volume_t),
                  family="binomial",
                  data=cactus)

cactus$Flow <- cactus$Goodbuds_t > 0 #blank column 
for(i in 1:8189){
  if(cactus$TotFlowerbuds_t[i] == NA){
    cactus$Flow[i] <- sum(cactus$ABFlowerbuds_t[i]+cactus$Goodbuds_t[i])
  }
  else if(cactus$TotFlowerbuds_t[i] == 0){
    cactus$Flow[i] <- 0
  }
  else if(cactus$TotFlowerbuds_t[i] >0){
    cactus$Flow[i] <- 1
  }
}

cactus$Flow
ggplot(cactus, aes(y=Flow, x=log(volume_t))) +  geom_point() +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE)

plot(log(cactus$volume_t),cactus$Flow)
```

#### Growth Model
```{r}
gro_model_ant_interact <- glm(log(volume_t1) ~ 0 + log(volume_t) * as.factor(bin1),
                  data=cactus)

ggplot(cactus, aes(y=log(volume_t1), x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', se = FALSE) 
```
## Ant Transition Model
```{r}
ant_trans_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1),family = binomial,data=cactus)
ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

```{r}
ant_trans_size_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) + log(volume_t),family = binomial,data=cactus)
ant_trans_size_model2 <- glm(as.factor(bin2) ~ 0 + as.factor(bin1) * log(volume_t),family = binomial,data=cactus)
#AIC(ant_trans_model)
#AIC(ant_trans_size_model)
#AIC(ant_trans_size_model2)
##Therefore we can see that the interaction transition model is the best. And that Size is an important factor.
summary(ant_trans_size_model2)
ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1)*log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
#model.matrix(ant_trans_size_model2)
```

## Multinomial Models

```{r}
#Will be ant vs size
x1 <- factor(as.factor(cactus$bin1), levels = c("5","1","2","3","4"))
m1 <- multinom(as.factor(cactus$bin1) ~ 0+ 1, data = cactus)
m2 <- multinom(x1 ~ 1, data = cactus)
z1<- summary(m1)$coefficients/summary(m1)$standard.errors
z2<- summary(m2)$coefficients/summary(m2)$standard.errors
summary(m1)
summary(m2)
#coef(m1)
#coef(m2)
p1 <- (1 - pnorm(abs(z1), 0, 1)) * 2
p2 <- (1 - pnorm(abs(z2), 0, 1)) * 2
ggplot(cactus, aes(x = cactus$bin2, y = cactus$bin1, colour = cactus$bin1)) + geom_line() 
#model.matrix(m1)s
```

```{r}
#Will be ant vs ant
m2 <- multinom(as.factor(bin1) ~ 0+as.factor(bin2), data = cactus)
z<- summary(m2)$coefficients/summary(m2)$standard.errors
#summary(m2)
p <- (1 - pnorm(abs(z), 0, 1)) * 2
ggplot(cactus, aes(x = bin2, y = bin1, colour = bin1)) + geom_line()
#model.matrix(m2)
```