### 2021-02-08 Stan/Stan Mix Models.Rmd

---
title: "Mixed Models Stan"
author: "Ali Campbell"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
library(rstanarm)
library(brms)  # for models
library(tidyverse)
library(rstan)
library(StanHeaders)
library(shinystan)
library(bayesplot)
library(devtools)
library(tidyr)
library(lattice)
library(lme4)
library(dplyr)
library(tidyverse)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
library("posterior")
#library(sjPlot)
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = parallel::detectCores())
setwd("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/ant_cactus_demography/Stan")
```

#### Data

```{r}
data <- cactus[ ,c("Plot","Year_t","Survival_t1","ant_t","ant_t1","volume_t","volume_t1")]
data <- na.omit(data)
data$ant <- as.integer(data$ant_t)
## add columns for each ant classification so they can be binary
data$crem <- 0 ##2
data$liom <- 0 ##3
data$other <- 0 ##1
data$vacant <- 0 ##4
for(i in 1:nrow(data)){
  if(data$ant[i] == 1){data$other[i] <- 1}
  if(data$ant[i] == 2){data$crem[i] <- 1}
  if(data$ant[i] == 3){data$liom[i] <- 1}
  if(data$ant[i] == 4){data$vacant[i] <- 1}
}
```

#### Mixed models
```{r growth}
## Write in Data
vol <- log(data$volume_t)
vol1 <- log(data$volume_t1)
ant <- as.numeric(data$ant_t)
plot <- as.factor(data$Plot)
year <- data$Year_t
y <- log(data$Survival_t1)
N <- length(data$volume_t)
K <- unique(data$ant_t)
J <- unique(data$Plot)
L <- unique(data$Year_t)
## Now make it a dataframe
stan_data <- list(N = N, vol = vol, vol1 = vol1, y = y, ant = ant, K = K, J = J, L = L)
write("// Stan model for simple linear regression

data {
  int <lower = 0> N; // number of observations
  int <lower = 0> K; // number of ant species
  real < lower = 0 > vol[N]; // volume t
  real < lower = 0 > vol1[N]; // volume t1
  int < lower = 0 > ant[K]; // ants
  int <lower=0,upper=1> y[N]; // outcome vector
  real rt[N];
  int<lower=1> J; //number of plots
  int<lower=1> L; //number of years
  int<lower=1, upper=J> plot[N]; // plot
  int<lower=1, upper=L> year[N]; // year
}

parameters {
  real alpha; // intercept
  vector[K] beta; // coefficients for ants
  vector[J] u; //subject intercepts
  vector[L] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}

transformed parameters {
  real mu;
}

model {
  // Priors
  alpha ~ normal(0,100);
  beta ~ normal(0,100); 
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  //sigma ~ exponential(0,0);
  // The Model
  for (i in 1:N){
    mu = beta[1] + u[plot[i]] + w[year[i]] //+ beta[2] * vol[i] + beta[3]*ant[i];
    rt[i] ~ lognormal(mu, sigma_e);
  }
}",

"grow_mixed1.stan")

stan_model1 <- "grow_mixed1.stan"
## Now we fit the model
fit <- stan(file = stan_model1, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit
```

```{r}
## Write in Data
vol <- log(data$volume_t)
vol1 <- log(data$volume_t1)
ant <- as.numeric(data$ant_t)
plot <- as.factor(data$Plot)
year <- data$Year_t
y <- log(data$Survival_t1)
N <- length(data$volume_t)
K <- as.factor(unique(data$ant_t))
J <- unique(as.factor(data$Plot))
L <- unique(data$Year_t)
## Now make it a dataframe
stan_data <- list(N = N, vol = vol, vol1 = vol1, y = y, ant = ant, K = K, J = J, L = L)
write("// Stan model for simple mixed model growth regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> K; // number of ant species
  real < lower = log(0) > vol[N]; // volume t
  int < lower = 1 > ant[K]; // ants
  real rt[N]; // response variable
  int < lower=1 > J; //number of plots
  int < lower=1 > L; //number of years
  int < lower=1, upper=J > plot[N]; // plot
  int < lower=1, upper=L > year[N]; // year
}

parameters {
  real alpha; // intercept
  vector[N] beta; // coefficients for ants
  vector[J] u; //subject intercepts
  vector[L] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}

transformed parameters {
  real mu[N];
  for (i in 1:N)
    mu[i] = beta[1] + u[plot[i]] + w[year[i]]+ beta[2] * vol[i] + beta[3]*ant[i];
}

model {
  // Priors
  alpha ~ normal(0,100);
  beta ~ normal(0,100); 
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  //sigma ~ exponential(0,0);
  // The Model
  for(i in 1:N)
    rt[i] ~ lognormal(mu[i], sigma);
}",

"grow_mixed1.stan")

stan_model1 <- "grow_mixed1.stan"
## Now we fit the model
fit <- stan(file = stan_model1, data = stan_data, warmup = 500, iter = 1000, chains = 4, cores = 2, thin = 1)
fit
```