### 2021-03-02 Stan/Stan Mix Models.Rmd
Ant Vital Rate Models (Stan)

---
title: "Mixed Models Stan"
author: "Ali Campbell"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
library(rstanarm)
library(brms)  # for models
library(tidyverse)
library(rstan)
library(StanHeaders)
library(shinystan)
library(bayesplot)
library(devtools)
library(tidyr)
library(lattice)
library(lme4)
library(dplyr)
library(tidyverse)
library(nnet)
library(fixest)
library(mlogit)
library(dfidx)
library(effects)
library(bbmle)
library(magrittr)
library(Gmisc)
library(RColorBrewer)
library(ggeffects)
library(heplots)
library("posterior")
#library(sjPlot)
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = parallel::detectCores())
setwd("/Users/alicampbell/Box Sync/Grad/Cactus/Shared Code/ant_cactus_demography/Stan")
```

#### Data

```{r}
data <- cactus[ ,c("Plot","Year_t","Survival_t1","ant_t","ant_t1","volume_t","volume_t1","repro_state_t1")]
data <- na.omit(data)
data$ant <- as.integer(data$ant_t)
## add columns for each ant classification so they can be binary
data$crem <- 0 ##2
data$liom <- 0 ##3
data$other <- 0 ##1
data$vacant <- 0 ##4
for(i in 1:nrow(data)){
  if(data$ant[i] == 1){data$other[i] <- 1}
  if(data$ant[i] == 2){data$crem[i] <- 1}
  if(data$ant[i] == 3){data$liom[i] <- 1}
  if(data$ant[i] == 4){data$vacant[i] <- 1}
}
data$Year_t <- as.factor(data$Year_t)
data$year <- as.integer(data$Year_t)
data$Plot <- as.factor(data$Plot)
data$plot <- as.integer(data$Plot)

flower <- cactus[ , c("TotFlowerbuds_t","ABFlowerbuds_t","Goodbuds_t","ant_t", "repro_state_t1","volume_t","volume_t1", "Year_t", "Plot")]
flower <- na.omit(flower)
flower$ant <- as.integer(flower$ant_t)
flower$Year_t <- as.factor(flower$Year_t)
flower$year <- as.integer(flower$Year_t)
flower$Plot <- as.factor(flower$Plot)
flower$plot <- as.integer(flower$Plot)
```

#### Growth
```{r Growth Mix Ant Model}
vol = data$volume_t
y = data$volume_t1
N = nrow(data)
N_ant = 4
ant = data$ant
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
                  ) 
## Write the model in stan
#note that we are using flat priors here which are non restrictive
write("// Stan model for simple growth regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  vector[N] y; // size_t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}

transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}

model {
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
  beta1 ~ normal(0,100); // slope distribution
  for(i in 1:N){
    y[i] ~ lognormal(mu[i], sigma);
  }
}",

"grow_ant_mix_lm1.stan")
## Check that you wrote a file
stanc("grow_ant_mix_lm1.stan")
## Save the file path
stanny <- "grow_ant_mix_lm1.stan"
## Now we fit the model
fit_grow_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5000, iter = 10000, chains = 1, cores = 2, thin = 1)
fit_grow_mix_ant
#class(fit)
posterior_grow_mix_ant <- as.data.frame(fit_grow_mix_ant)
#print(colnames(posterior_grow_mix_ant))
lm_alpha_other <- summary(grow_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(grow_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(grow_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(grow_model_ant_interact)$coeff[5]
lm_beta_other <- summary(grow_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(grow_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(grow_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(grow_model_ant_interact)$coeff[8]
lm_sigma <- sigma(grow_model_ant_interact)  # the residual error

par(mfrow = c(2,4))
#Plot Intercepts
plot(density(posterior_grow_mix_ant$`beta0[1]`), main = "Intercept Other")
abline(v = lm_alpha_other, col = 4, lty = 2)
plot(density(posterior_grow_mix_ant$`beta0[2]`), main = "Intercept Crem")
abline(v = lm_alpha_crem, col = 4, lty = 2)
plot(density(posterior_grow_mix_ant$`beta0[3]`), main = "Intercept Liom")
abline(v = lm_alpha_liom, col = 4, lty = 2)
plot(density(posterior_grow_mix_ant$`beta0[4]`), main = "Intercept Vacant")
abline(v = lm_alpha_vacant, col = 4, lty = 2)
#Plot Slopes
plot(density(posterior_grow_mix_ant$`beta1[1]`), main = "Intercept Other")
abline(v = lm_beta_other, col = 4, lty = 2)
plot(density(posterior_grow_mix_ant$`beta1[2]`), main = "Intercept Crem")
abline(v = lm_beta_crem, col = 4, lty = 2)
plot(density(posterior_grow_mix_ant$`beta1[3]`), main = "Intercept Liom")
abline(v = lm_beta_liom, col = 4, lty = 2)
plot(density(posterior_grow_mix_ant$`beta1[4]`), main = "Intercept Vacant")
abline(v = lm_beta_vacant, col = 4, lty = 2)
#Sigma
plot(density(posterior_grow_mix_ant$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)

```

####Survival
```{r survival}
vol = data$volume_t
y = data$Survival_t1
N = nrow(data)
N_ant = 4
ant = data$ant
N_Year <- max(data$year)
N_Plot <- max(data$plot)
plot <- data$plot
year <- data$year
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
                  )

write("// Stan model for simple ant survival regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0, upper = 1> y[N]; // survival in year t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}

parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}

transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}

model {
//Priors
 u ~ normal(0, sigma_u); // plot random effects
 w ~ normal(0, sigma_w); // year random effects
 beta0 ~ normal(0,100); // intercept distribution
 beta1 ~ normal(0,100); // slope distribution
 //Model
 for(i in 1:N){
 y[i] ~ bernoulli_logit(mu[i]);
 }
}
",

"surv_ant_mix_lm1.stan")
## Check that you wrote a file
stanc("surv_ant_mix_lm1.stan")
## Save the file path
stanny <- "surv_ant_mix_lm1.stan"
## Now we fit the model
fit_surv_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_surv_mix_ant

#class(fit)
posterior_surv_mix_ant <- as.data.frame(fit_surv_mix_ant)
print(colnames(posterior_surv_mix_ant))
lm_alpha_other <- summary(surv_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(surv_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(surv_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(surv_model_ant_interact)$coeff[5]
lm_beta_other <- summary(surv_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(surv_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(surv_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(surv_model_ant_interact)$coeff[8]
lm_sigma <- sigma(surv_model_ant_interact)  # the residual error

par(mfrow = c(2,4))
#Plot Intercepts
plot(density(posterior_surv_mix_ant$`beta0[1]`), main = "Intercept Other")
abline(v = lm_alpha_other, col = 4, lty = 2)
plot(density(posterior_surv_mix_ant$`beta0[2]`), main = "Intercept Crem")
abline(v = lm_alpha_crem, col = 4, lty = 2)
plot(density(posterior_surv_mix_ant$`beta0[3]`), main = "Intercept Liom")
abline(v = lm_alpha_liom, col = 4, lty = 2)
plot(density(posterior_surv_mix_ant$`beta0[4]`), main = "Intercept Vacant")
abline(v = lm_alpha_vacant, col = 4, lty = 2)
#Plot Slopes
plot(density(posterior_surv_mix_ant$`beta1[1]`), main = "Intercept Other")
abline(v = lm_beta_other, col = 4, lty = 2)
plot(density(posterior_surv_mix_ant$`beta1[2]`), main = "Intercept Crem")
abline(v = lm_beta_crem, col = 4, lty = 2)
plot(density(posterior_surv_mix_ant$`beta1[3]`), main = "Intercept Liom")
abline(v = lm_beta_liom, col = 4, lty = 2)
plot(density(posterior_surv_mix_ant$`beta1[4]`), main = "Intercept Vacant")
abline(v = lm_beta_vacant, col = 4, lty = 2)
#Sigma
plot(density(posterior_surv_mix_ant$sigma), main = "Sigma")
abline(v = lm_sigma, col = 4, lty = 2)
```

####Total Flowers

```{r total flowers}
y <- flower$TotFlowerbuds_t
vol = flower$volume_t
N = nrow(flower)
N_ant = 4
ant = flower$ant
N_Year <- max(flower$year)
N_Plot <- max(flower$plot)
plot <- flower$plot
year <- flower$year
stan_data <- list(N = N, ## number of observations
                  vol = vol, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
                  )

write("// Stan model for simple total flowers regression

data {
 int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol;	//size_t
  int <lower = 0> y[N]; // survival in year t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}

parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}

transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol[i] + u[plot[i]] + w[year[i]];
  }
}
model {
 // Model
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
  beta1 ~ normal(0,100); // slope distribution
  for(i in 1:N){
    y[i] ~ poisson(mu[i]);
  }
}
",

"flower_ant_mix_lm1.stan")
## Check that you wrote a file
stanc("flower_ant_mix_lm1.stan")
## Save the file path
stanny <- "flower_ant_mix_lm1.stan"
## Now we fit the model
fit_flow_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_flow_mix_ant

posterior_flow_ant <- as.data.frame(fit_flow_mix_ant)
print(colnames(posterior_flow_mix_ant))

lm_alpha_other <- summary(flower_model_ant)$coeff[1]  # the intercept
lm_alpha_crem <- summary(flower_model_ant)$coeff[3]
lm_alpha_liom <- summary(flower_model_ant)$coeff[4]
lm_alpha_vacant <- summary(flower_model_ant)$coeff[5]
lm_beta_other <- summary(flower_model_ant)$coeff[2]  # the slope
lm_beta_crem <- summary(flower_model_ant)$coeff[6]
lm_beta_liom <- summary(flower_model_ant)$coeff[7]
lm_beta_vacant <- summary(flower_model_ant)$coeff[8]

par(mfrow = c(2,4))
#Plot Intercepts
plot(density(posterior_flow_mix_ant$`beta0[1]`), main = "Intercept Other")
abline(v = lm_alpha_other, col = 4, lty = 2)
plot(density(posterior_flow_mix_ant$`beta0[2]`), main = "Intercept Crem")
abline(v = lm_alpha_crem, col = 4, lty = 2)
plot(density(posterior_flow_mix_ant$`beta0[3]`), main = "Intercept Liom")
abline(v = lm_alpha_liom, col = 4, lty = 2)
plot(density(posterior_flow_mix_ant$`beta0[4]`), main = "Intercept Vacant")
abline(v = lm_alpha_vacant, col = 4, lty = 2)
#Plot Slopes
plot(density(posterior_flow_mix_ant$`beta1[1]`), main = "Intercept Other")
abline(v = lm_beta_other, col = 4, lty = 2)
plot(density(posterior_flow_mix_ant$`beta1[2]`), main = "Intercept Crem")
abline(v = lm_beta_crem, col = 4, lty = 2)
plot(density(posterior_flow_mix_ant$`beta1[3]`), main = "Intercept Liom")
abline(v = lm_beta_liom, col = 4, lty = 2)
plot(density(posterior_flow_mix_ant$`beta1[4]`), main = "Intercept Vacant")
abline(v = lm_beta_vacant, col = 4, lty = 2)
```

####Viability Model

```{r Viability model}
good <- flower$Goodbuds_t
abort <- flower$ABFlowerbuds_t
tot <- flower$TotFlowerbuds_t
vol = log(flower$volume_t)
N = nrow(flower)
N_ant = 4
ant = flower$ant
N_Year <- max(flower$year)
N_Plot <- max(flower$plot)
plot <- flower$plot
year <- flower$year
stan_data <- list(N = N, ## number of observations = trials
                  N_ant = N_ant,
                  ant = ant,
                  good = good, ## good buds = successes
                  abort = abort, ## aborted buds data
                  tot = tot, ## number of trials
                  vol = vol,
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
)

write("// Stan model for simple total flowers regression
data {
 int < lower = 1 > N; // Sample size
 int <lower = 0> good[N]; // Number of viable seeds
 int <lower = 1> N_ant; // number of ant states
 int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
 int<lower=0> tot[N]; // number of trials
 int<lower=0> abort[N];
 vector[N] vol;	//size_t
 int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}
parameters {
  real <lower = 0, upper = 1> theta[N]; // probability of viability for each bud
  //
  real<lower=0,upper=1> v0;
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}
transformed parameters{
  real<lower=0,upper=1> predV[N]; // proportion viable for each plant?
  // Prediction for seed viability
  for(i in 1:N){
    predV[i] = v0 * pow(theta[ant[i]],good[i]) * (1-theta[ant[i]])^abort[i] + u[plot[i]] + w[year[i]];
  }
}
model {
 // Priors
 v0  ~ beta(10,1);  // intercept viability model
  u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  theta ~ normal(0,100); // intercept distribution

  good ~ binomial(tot, predV);
}
", "viab_mix_ant.stan")
## Check that you wrote a file
stanc("viab_mix_ant.stan")
## Save the file path
stanny <- "viab_mix_ant.stan"
## Now we fit the model
fit_viab_mix_ant <- stan(file = stanny, data = stan_data, warmup = 5, iter = 10, chains = 1, cores = 2, thin = 1)

posterior_viab_mix_ant <- as.data.frame(fit_viab_mix_ant)

lm_alpha_other <- summary(viab_model_ant_interact)$coeff[1]  # the intercept
lm_alpha_crem <- summary(viab_model_ant_interact)$coeff[3]
lm_alpha_liom <- summary(viab_model_ant_interact)$coeff[4]
lm_alpha_vacant <- summary(viab_model_ant_interact)$coeff[5]
lm_beta_other <- summary(viab_model_ant_interact)$coeff[2]  # the slope
lm_beta_crem <- summary(viab_model_ant_interact)$coeff[6]
lm_beta_liom <- summary(viab_model_ant_interact)$coeff[7]
lm_beta_vacant <- summary(viab_model_ant_interact)$coeff[8]

par(mfrow = c(1,2))
#Plot Slopes
plot(density(posterior_viab_ant$`theta[1]`), main = "Viability Slope Distributions", col = "black", lwd = 2, xlim = c(-1,3), ylim =  c(0,2))
abline(v = lm_alpha_other, col = "black", lty = 2)
lines(density(posterior_viab_ant$`theta[2]`),  col = "red")
abline(v = lm_alpha_crem, col = "red", lty = 2)
lines(density(posterior_viab_ant$`theta[3]`),col = "blue", lwd = 2)
abline(v = lm_alpha_liom, col = "blue", lty = 2)
lines(density(posterior_viab_ant$`theta[4]`), col = "pink", lwd = 2)
abline(v = lm_alpha_vacant, col = "pink", lty = 2)
legend("topleft", legend = c("other","Crem.", "Liom.","Vac"), col = c("black","red","blue","pink"), lty = 1)
```

####Repro Model
```{r}
vol1 <- flower$volume_t1
y = flower$repro_state_t1
N = nrow(flower)
N_ant = 4
ant = flower$ant
N_Year <- max(flower$year)
N_Plot <- max(flower$plot)
plot <- flower$plot
year <- flower$year
stan_data <- list(N = N, ## number of observations
                  vol1 = vol1, ## predictors volume
                  y = y, ## response volume next year
                  ant = ant,## predictors ants
                  N_ant = N_ant, ## number of ant states
                  N_Year = N_Year, ## number of years
                  N_Plot = N_Plot, ## number of plots
                  plot = plot, ## predictor plots
                  year = year ## predictor years
                  )

write("// Stan model for simple total flowers regression

data {
  int <lower = 1> N; // number of observations
  int <lower = 1> N_ant; // number of ant states
  int <lower = 1, upper = N_ant> ant[N]; // the list of ant species 
  vector[N] vol1;	//size_t
  int <lower = 0, upper = 1> y[N]; // survival in year t1
  int<lower=1> N_Year; //number of plots
  int<lower=1> N_Plot; //number of years
  int<lower=1, upper=N_Plot> plot[N]; // plot
  int<lower=1, upper=N_Year> year[N]; // year
}

parameters {
  vector[N_ant] beta0; //intercept, unique to ant sp
  vector[N_ant] beta1; //slope, unique to ant sp
  vector[N_Plot] u; //subject intercepts
  vector[N_Year] w; //item intercepts
  real < lower = 0 > sigma; // Error SD
  real < lower = 0 > sigma_u; // plot SD
  real < lower = 0 > sigma_w; // year SD
}

transformed parameters{
  vector[N] mu; //linear predictor for the mean
  for(i in 1:N){
    mu[i] = beta0[ant[i]] + beta1[ant[i]] * vol1[i] + u[plot[i]] + w[year[i]];
  }
}
model {
 // Priors
 `u ~ normal(0, sigma_u); // plot random effects
  w ~ normal(0, sigma_w); // year random effects
  beta0 ~ normal(0,100); // intercept distribution
  beta1 ~ normal(0,100); // slope distribution
 // Model
 for(i in 1:N){
  y[i] ~ bernoulli_logit(mu[i]);
 }
}
",

"repro_lm1.stan")

## Check that you wrote a file
stanc("repro_ant_mix_lm1.stan")
## Save the file path
stanny <- "repro_ant_mix_lm1.stan"
## Now we fit the model
fit_repro_mix_ant <- stan(file = stanny, data = stan_data, warmup = 500, iter = 1000, chains = 1, cores = 2, thin = 1)
fit_repro_mix_ant

posterior_repro_mix_ant <- as.data.frame(fit_repro_mix_ant)
print(colnames(posterior_repro_mix_ant))

lm_alpha_other <- summary(model3)$coeff[1]  # the intercept
lm_alpha_crem <- summary(model3)$coeff[3]
lm_alpha_liom <- summary(model3)$coeff[4]
lm_alpha_vacant <- summary(model3)$coeff[5]
lm_beta_other <- summary(model3)$coeff[2]  # the slope
lm_beta_crem <- summary(model3)$coeff[6]
lm_beta_liom <- summary(model3)$coeff[7]
lm_beta_vacant <- summary(model3)$coeff[8]

par(mfrow = c(2,4))
#Plot Intercepts
plot(density(posterior_repro_mix_ant$`beta0[1]`), main = "Intercept Other")
abline(v = lm_alpha_other, col = 4, lty = 2)
plot(density(posterior_repro_mix_ant$`beta0[2]`), main = "Intercept Crem")
abline(v = lm_alpha_crem, col = 4, lty = 2)
plot(density(posterior_repro_mix_ant$`beta0[3]`), main = "Intercept Liom")
abline(v = lm_alpha_liom, col = 4, lty = 2)
plot(density(posterior_repro_mix_ant$`beta0[4]`), main = "Intercept Vacant")
abline(v = lm_alpha_vacant, col = 4, lty = 2)
#Plot Slopes
plot(density(posterior_repro_mix_ant$`beta1[1]`), main = "Intercept Other")
abline(v = lm_beta_other, col = 4, lty = 2)
plot(density(posterior_repro_mix_ant$`beta1[2]`), main = "Intercept Crem")
abline(v = lm_beta_crem, col = 4, lty = 2)
plot(density(posterior_repro_mix_ant$`beta1[3]`), main = "Intercept Liom")
abline(v = lm_beta_liom, col = 4, lty = 2)
plot(density(posterior_repro_mix_ant$`beta1[4]`), main = "Intercept Vacant")
abline(v = lm_beta_vacant, col = 4, lty = 2)
```